Nanophotonics with optical phonons in two dimensions

### 2018-11-07 paper.tex
need to fill out one more section

@@

To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges. Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in [Needs reference!], one finds that the RPA dielectric function is given by:
\begin{equation}
\Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q}L}}{\omega^2_{\mathbf{q}TO}-\omega^2}|\hat{q}\cdot\langle 1_{\mathbf{q}L}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2.
\end{equation}
In this expression, $\omega_L$ and been re-named as $\omega_{TO}$, because in the absence of LO-TO splitting, they are degenerate. Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.

As is well known, the zeros of the RPA dielectric function correspond to longitudinal electromagnetic waves. Thus, the condition for the zeros of the dielectric function correspond to
\begin{equation}
\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},T} = \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2
\end{equation}

@@

@@

%\omega = \sqrt{\omega_{\mathrm{TO}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.

@@

%Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{TO}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
%In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{TO}}-\omega^2}$. It is interesting to see that in the 2D limit, the effect of the screening inside the material altogether vanishes. A similar phenomenon is known in the electrodynamics of plasmons in 2D materials, where the dielectric screening is known to be entirely provided by the environment and is equal to $\epsilon_{\mathrm{env}}$. A useful result which follows is that we may also derive another important figure of merit, the $p$-polarized reflectivity of a 2D polar dielectric, which is related to the local density of states of the polar dielectric, which quantifies the strength of coupling to external probes such as a nano-antenna or a quantum emitter.

@@

% It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case. As a recently discovered example of such a phenomenon, it was found that the LO-TO splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector. This would then suggest that a description of phonon polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short. Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions. For example, it was found in several experiments that the damping rate of the TO phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
%%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the TO phonon frequency.
%%Near the TO phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.

>>>>

In this Letter, we ...

@@

To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges. Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in [Needs reference!], one finds that the RPA dielectric function is given by:
\begin{equation}
\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}
\end{equation}
In this expression, $\omega_L$ and been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate. Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.

As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves. Thus, the condition for the zeros of the dielectric function correspond to
\begin{align}
\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
&= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
\end{align}
where $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates. This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon. In particular, it was shown in [Sohier2017] that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting given by Equation (8) of the current manuscript. \textit{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}

We now re-write $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit. From Equation (8), we may write Equation (7) as
\begin{equation}
\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{\omega^2_{\mathbf{q}\mathrm{LO}}-\omega^2_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2}\right).
\end{equation}
Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$. We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases [ref], the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
\begin{equation}
\omega_{\mathbf{q},LO} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
\end{equation}
where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_g$ is the group velocity of the LO mode, and $\omega_{\Gamma \mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point. Then, the universal form of the dispersion relation 
\begin{equation}
\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right).
\end{equation}
where we have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription. Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$.

@@

We now move to analyze key figures of merit of 2D LO phonons.

@@

%\omega = \sqrt{\omega_{\mathrm{\mathrm{TO}}}^2+\frac{n_{\mathrm{s}}Q_{\mathrm{eff}}^{2}}{2M_{\mathrm{eff}}\epsilon_0\epsilon_{\mathrm{env}}}q}.

@@

%Taking Equation~\eqref{eq:lorentzoscillator} for the dielectric function in bulk, we have that (in the absence of losses) $\epsilon_{\mathrm{env}} + qd\epsilon_{\infty} + q\frac{n_sQ_{\mathrm{eff}}^{2}}{2\epsilon_0 M_{\mathrm{eff}}}\frac{1}{\omega_{\mathrm{\mathrm{TO}}}^2-\omega^2} = 0$, where we have identified $V = 2Ad$.
%In the limit of small $d$, i.e., $qd \ll 1$, we recover Equations~\eqref{eq:2ddispersion} and \eqref{eq:2dphononpolaritondispersion} with $\Pi = \frac{n_sQ_{\mathrm{eff}}^{2}}{\epsilon_0 M}\frac{1}{\omega^2_{\mathrm{\mathrm{TO}}}-\omega^2}$. It is interesting to see that in the 2D limit, the effect of the screening inside the material altogether vanishes. A similar phenomenon is known in the electrodynamics of plasmons in 2D materials, where the dielectric screening is known to be entirely provided by the environment and is equal to $\epsilon_{\mathrm{env}}$. A useful result which follows is that we may also derive another important figure of merit, the $p$-polarized reflectivity of a 2D polar dielectric, which is related to the local density of states of the polar dielectric, which quantifies the strength of coupling to external probes such as a nano-antenna or a quantum emitter.

@@

% It is also quite possible that they altogether vanish in the 2D case as opposed to the 3D case. As a recently discovered example of such a phenomenon, it was found that the LO-\mathrm{TO} splitting of some 2D polar dielectrics, such as hBN \cite{sohier2017breakdown} altogether vanishes, and only returns to its bulk value after a small but finite wavevector. This would then suggest that a description of phonon polaritons necessitates consideration of the spatially nonlocal dielectric function as in Equation (12), and that these polaritons moreover may only exist when their wavelengths are sufficiently short. Another important change, beyond the theory presented in this work is that the losses may also change drastically in the reduction of dimensions. For example, it was found in several experiments that the damping rate of the \mathrm{TO} phonon in hBN nearly doubles upon going from bulk to 2D\cite{gorbachev2011hunting,tran2016quantum}, which may strongly suppress the ability to image the polariton with near-field probes.
%%This is due to the small thickness where the dispersion of the phonon polariton is quite flat, and very high wavevector modes, corresponding to the maximum wavevector detectable by a near-field tip, occur at frequencies very close to the \mathrm{TO} phonon frequency.
%%Near the \mathrm{TO} phonon-frequency however, the ratio $\Im\Pi/\Re\Pi$ is the highest, corresponding to very low quality-factor modes.
